Composable Generalized Consensus

### 2016-01-07 paper/abstract4.tex
Working on the abstract: grammar and emphasis on the problem.

State-Machine Replication (SMR) algorithms are notoriously hard to understand
and devising new and more performant SMR algorithms is costly. To lower this
cost, we present a framework for building SMR algorithms out of independent
and reusable modules. A module is an SMR algorithm that can abort and pass
the baton to another module, allowing to dynamically switch between modules.
Our composition theorem ensures that the composition of any number of modules
satisfying Composable Generalized Consensus (CGC), our correctness condition for
modules, is a correct SMR algorithm.

To help design new modules in practice, we propose TLA$^+$ specifications
of CGC and of two abstract types of modules, fast rounds and conservative
rounds, each providing a template to devise new modules with the help of the
TLA$^+$ tools. We show that several well-known SMR algorithms can be modified to
implement fast or conservative rounds, therefore enabling their composition to
obtain a single versatile SMR algorithm.

>>>>

State-Machine Replication (SMR) algorithms are notoriously hard to understand, and one can trust their correctness only by a costly formal specification and analysis.
Moreover, it is often impossible to reuse the results of formal analyses because
existing SMR algorithms are not modular.
The design of new SMR algorithms is therefore expensive.

To lower the cost of designing new SMR algorithms, we present a framework for composing independent and reusable SMR modules. 
An SMR module is an SMR algorithm augmented with a composition interface. 
In the composition of two modules, the first module starts executing and can abort its execution and pass the baton to the second module at any time.
Our composition theorem guarantees that the result of composing two SMR modules is again an SMR module and a correct SMR algorithm.
Module composition can be dynamic; therefore, our framework allows building SMR algorithms that choose and switch at runtime between a set of modules, and that can be augmented with new modules at any time.

To help design new modules in practice, we propose several TLA$^+$ specifications that each
provide a template to devise new modules with the help of the TLA$^+$ tools.
Using our specifications, we demonstrate that several well-known SMR algorithms can be made modular and composed to form a versatile SMR algorithm.
